---
title: "The Solution to Assignment 1"
author: "Xiaojun Sun"
date: "Sunday, June 29, 2014"
output:
  html_document:
    theme: united
---

Dear all WISE R Club members. This is the solution to our first assignment. I hope you guys be more familiar with R after reading the book and finishing assignment 1. 

```{r initial, echo=FALSE}
setwd("E:/RSpace/R Club/assignment1")
```

### Question 1
First of all, we load 
```{r loaddata,cache=TRUE}
grades <- read.table("grades.txt", header=TRUE,stringsAsFactors=FALSE)
# `stringsAsFactors=FALSE` prevent the `read.table` function from treating the character name as factor
# `grades` is a data frame
```

After creating the dataframe you can see what's in it using the following functions.

```{r detail}
head(grades)
str(grades)
```


### Question 2
We plot the histogram of the grade of math.
```{r hist,fig.width=8,fig.cap="The histgram of math grade"}
hist(grades$Math,col="lightblue",main="The histogram of the grade of math",xlab="The grade of math")
```

### Question 3
Now we plot the scatterplot of `Math vs. Chinese` and `Math vs. English` in one long graph.
```{r xyplot, fig.width=8, fig.height=12,fig.cap="The Scatterplots"}
opar <- par(no.readonly=TRUE)
par(mfcol=c(2,1))
par(lwd=1.5)
attach(grades)
plot(Math,Chinese,col="red",main="Scatterplot of Math vs. Chinese")
abline(lm(Math~Chinese))
plot(Math,English,col="blue",main="Scatterplot of Math vs. English")
abline(lm(Math~English))
detach(grades)
par(opar)
```

### Question 4
To see how many students have passed this exam.
```{r passexam}
pass <- subset(grades,Math>=60 & Chinese>=60 & English>=60, select=name)
dim(pass)[1]
```

So in total `r dim(pass)[1]` students have passed the exam.

### Question 5
To find the out the standing student.
```{r outstanding}
outstand <- subset(grades,Math>=90 & Chinese>=90 & English>=90, select=name)
outstand
```

So the headmaster should award `r outstand`. Congratulations!

### Note
The solution is not unique.

Someone told me that Chinese characters cann't display normally in R. I have to say R is friendly enough to Chinese character. If there is something wrong when you try to display Chinese character, it's highly probably that you made a mistakes when loading the data, that is, you didn't specify `stringsAsFactors=FALSE`.

To display that R is friendly enough to Chinese character I can give you an example. 

```{r chinese}
变量1 <- 2
变量2 <- 3
变量1+变量2
变量1*变量2
```

As you see, the Chinese characters can even be the variable name. How amazing is it!

If you have more questions about this solution, you can send an email to me or ask me in the next meeting.

Thank you!

__The end__.
