---
title: "The Solution to Assignment 2"
author: "Xiaojun Sun"
date: "Saturday, July 12, 2014"
runtime: shiny
output:
  html_document:
    fig_height: 7
    fig_width: 10
    highlight: textmate
    theme: cerulean
---

Dear all WISE R Club members. :smile: Here comes the solution to assignment2. Because assignment 2 is more difficult than assignment 1, I advise you to read this solution manual more carefully to make sure you understand every single code. 

```{r initial, echo=FALSE}
setwd("E:/RSpace/R Club/assignment2")
```

**********************

### Question 1
I define the factorial function in this way.
```{r factorial}
Fac <- function(n) if (n==1|n==0) 1 else n*Fac(n-1)
```
Now test it!
```{r testFac}
c(Fac(0),Fac(3),Fac(5))
```

**********************

### Question 2
To start with, I read the `grades.txt` dataset to create data frame.
```{r loaddata,cache=TRUE}
grades <- read.table("grades.txt", header=TRUE,stringsAsFactors=FALSE)
```

#### Question 2.1
To plot the boxplot, we should reshape the data frame `grades` at first.

```{r melt}
library(reshape2)
gradex <- melt(grades,id="name")
```

We can see what is in the new data frame `gradex`.
```{r intoframe}
head(gradex)
str(gradex)
```

Then we plot the boxplot just like Figure 6.12.
```{r boxplot,fig.cap="the boxplot of test scores"}
boxplot(value ~ variable, data=gradex, col="lightblue",
        main="The boxplot of the test scores", 
        xlab="Subject", ylab="Test Scores")
```


#### Question 2.2
To calculate `Sum` and get the 80th, 60th, 40th and 20th percentile of it,
```{r getSum}
Sum <- rowSums(grades[,2:4])
x <- quantile(Sum, c(0.8,0.6,0.4,0.2))
```

To grade the students,
```{r grade}
grades <- cbind(grades, Sum)
grades$grade[Sum >= x[1]] <-"A+"
grades$grade[Sum < x[1] & Sum >= x[2]] <-"A"
grades$grade[Sum < x[2] & Sum >= x[3]] <-"B"
grades$grade[Sum < x[3] & Sum >= x[4]] <-"C"
grades$grade[Sum < x[4]] <-"D" 
str(grades)
```


#### Question 2.3

```{r Gradefunc}
Grade <- function(nm){
  names <- grades$name
  grades$grade[nm==names]
}
## see a few examples
c(Grade("ÍõÊéÆå"), Grade("Ò¦ÊËÄÜ"), Grade("ÏÄ¡¡Á¼"))
```


#### Question 2.4
Now I draw the pie chart of the grade.
```{r pie,fig.cap="the pie chart of grade"}
y <- table(grades$grade)
lb <- paste(names(y), " ", y, sep="")
pie(y, labels=lb, main= "Pie chart of the grade \n (with sample size)")
box()
```


#### Question 2.5
Now calculate the mean and standard error for each subject in each group.

```{r desc}
options(digits=2)
dstats <- function(x)(c(mean=mean(x), sd=sd(x)))
aggregate(grades[,2:4], by=list(grade=grades$grade), dstats)
```


**********************

### Question 3

#### Question 3.1
Regress `Math` on `Chinese` and `English`.
```{r regression}
fit<-lm(Math~Chinese+English,data=grades)
summary(fit)
par(mfrow=c(2,2))
plot(fit)
```

#### Question 3.2
Now I plot a histogram of the studentized residuals below and superimposes a normal curve, kernel density curve and rug plot.
```{r plotresidual}
z <- rstudent(fit)
hist(z, breaks=13, freq=FALSE,ylim=c(0,0.43), xlab="Studentized Residual",main="Distribution of Errors")
rug(jitter(z), col="brown")
curve(dnorm(x, mean=mean(z), sd=sd(z)), add=TRUE, col="blue", lwd=2)
lines(density(z)$x, density(z)$y, col="red", lwd=2, lty=2)
legend("topright",legend = c( "Normal Curve", "Kernel Density Curve"),lty=1:2, col=c("blue","red"), cex=1)
box()
```


<p>&nbsp;</p>
<p><a href="http://xiaojunsun.github.io/wise-r-club/">Return to the homepage.</a></p>
*****************
<p></a>WISE R Club project is proudly maintained by <a href="https://github.com/XiaojunSun">XiaojunSun</a>.</p>
